---
title: "CommunitySim"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CommunitySim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette, we simulate the dynamics of a community of bacteria after inoculation, and use various popular machine learning algorithms to predict the steady state abundance of the bacteria using their initial abundance. In the process, we demonstrate how to use CommunitySim package to simulate the parameters for community simulation, how to simulate a community, and how to read off the steady state abundance of members of the community.

```{r setup}
library(CommunitySim)
```

## Basic Functions

In this demonstration, we simulate a community of 8 possible species. 4 parameters are needed to describe and hence simulate the community: `alpha` is 1x8 vector of intrinsic growth rate of each species; `c0` is a 8x8 matrix of pairwise interaction intensity in the community, where `c0[i,j]` is the impact of species `j` on `i` on a per unit basis; `ck` is a list of 8 matrices, each with 8x8 entries, where `ck[[k]][i,j]` is the higher order impact from species `j` onto the pairwise interaction of `i` and `k`, felt by `k`; `init` is a 1x8 vector of initial abundance at the inoculation.

```{r}
N <- 8
params <- simulateParam(N)
print(diag(params$c0))
```

Notice the diagonal of `c0` are all set to `-0.5` to reflect the intra-species competition.

Then we simulate the dynamics of the abundance of species. A DataFrame of 8 columns (for each species) of abundance at each integration time points (default to 300 time points) is returned. A plot of all columns are also displayed.

```{r}
abundance <- growthFunction(params$N, params$alpha, params$c0, params$ck, params$init)
```

## Machine Learning Case Study

Suppose we know an environment has 8 possible bacterial species. When we take sample from the environment, it is not always the case for all 8 species to be sampled. In fact, there are 2^8 possibilities of how many of the 8 species are sampled. If we assume a fixed inoculation abundance for each species, we can generate 2^8 initial conditions.

```{r}
mask <- binaryInitState(N)
head(mask)

init <- t(t(mask) * params$init)
head(data.frame(init))
```
where each row correspond to an inoculation abundance when some species are present while the others are not.

Next, for each row - an initial abundance, we can run the community simulation of such inoculation.

```{r fig.show='hide'}
abun_list <- apply(init, 1, function(x) growthFunction(N, params$alpha, params$c0, params$ck, x))
```

We then use a function provided by CommunitySim to find the steady state abundance of each species for each initial state simulation.

```{r}
matrixToSS <- function(densityMatrix){
  s <- apply(densityMatrix[, 2:(N+1)], MARGIN = 2, FUN=findSteadyState)
  return(s)
}

SS <- sapply(abun_list, matrixToSS)
SS <- t(SS)
#If SS density is too small, set to 0
SS[which(SS < 0.001)] = 0
```
